The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-18 writing/draft3/draft15.tex
writing hardly

\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \cite{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \cite{bates2010lme4}. On the other hand, random effects are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation %(\citealp{KERYch9, KERYch12}).

@@

According to a shape of a histogram of the maize yield data we decided to log-transform the yield data before conducting our analysis. Hence, we used a log-linear functional form of our models as is a common practice in agricultural economics. Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model which minimized the value of the Akaike’s Information Criterion (AIC). We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors). We decided to not to include the number of heatwave days and average maximum daily temperature although including them in the fixed effects slightly improved the AIC. The reasons for not including them were their strong correlation with average seasonal temperature (the value of the correlation coefficient was $0.981$ for the correlation of the average seasonal temperature with the maximum daily temperature and $0.517$ the correlation of the average seasonal temperature with the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and the lack of significance of their individual t-tests (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our expectation. Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.

>>>>

\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \cite{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed effects parameters and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \cite{bates2010lme4}. On the other hand, random effects are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation \cite{KERYch9,KERYch12}.

@@

It was apparent from the histogram of the maize yield data that the distribution of maize yield was closer to a log-normal than to a normal distribution. Therefore, we opted for a log-linear functional form of our models as is a common practice in agricultural economics. Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure which minimized the Akaike's Information Criterion (AIC). The first step of the procedure was the estimation of the model with complete set of our weather measures in the fixed effects and with the random intercepts. After this we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors). We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects slightly improved the AIC. The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including them only improved AIC from $2080.877$ to $2076.439$ (these values differ from the AIC values presented
in table~\ref{ARMA} in the appendix because the final estimates compared in table~\ref{ARMA} were estimated using restricted maximum likelihood (REML), but the models were re-estimating using ML method for the comparison as REML comparisons are not meaningful). In addition, the signs of these variables were positive in contrary to our expectation. Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.
